---
title: "explore_forecast_ui"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(reticulate)
library(dplyr)
library(ggplot2)
library(DBI)
library(zoo)
library(dygraphs)
library(xts)

df_to_ts <- function(df){
  xts(
    x = df[,2:ncol(df)],
    order.by = df[,1]
    )
}
```

```{r}
# Connect to "jdbc:postgresql://postgres:5432/test"
con <- DBI::dbConnect(
  drv = RPostgres::Postgres(),
  host = "postgres",
  port=5432,
  dbname = "test",
  user = "test",
  password = "postgres"
  )
```

Check unique `connection_id`
```{sql connection=con, output.var='unique_connection_id'}
select distinct connection_id from whale_historical_ptu
```

Pick the first one for the experiment
```{r}
picked_connection_id <- unique_connection_id$connection_id[1]
picked_connection_id
```

Use such connection to see models associated
```{sql connection=con, output.var='model_df'}
select *
from whale_models
where connection_id = ?picked_connection_id
```

```{r}
head(model_df)
```
From these, pick the first model
```{r}
picked_model_id <- model_df$id[1]
picked_model_id
```

And use this model to get the list of available backcast
```{sql connection=con, output.var='backcast_df'}
select *
from whale_backcast wb
where model_id = ?picked_model_id
```

```{r}
head(backcast_df)
```
Of which we take the first one

```{r}
picked_backcast_id <- backcast_df$id[1]
```

----

We proceed to fetch the actuals using `picked_connection_id`
```{sql connection=con, output.var='actuals_fetched'}
select datetime, consumption
from whale_historical_ptu whp 
where connection_id = ?picked_connection_id
and consumption is not null
```

```{r}
head(actuals_fetched)
```

The backcast is stored in a csv; we fetch it and read it
```{r}
backcast_fetched <- readr::read_csv(glue::glue("/usr/local/remote_data/whale_backcast/{picked_backcast_id}.csv"))
```
Last, we visualize both together
```{r}
full_join(actuals_fetched, backcast_fetched, by = "datetime") %>% 
  df_to_ts() %>% 
  dygraph() %>% 
  dyRangeSelector()
```



